---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
dat<-read.csv("activity.csv",stringsAsFactors=F)
options(scipen=100,digits=4)
```



## What is mean total number of steps taken per day?
```{r}
x<-as.data.frame(tapply(dat$steps,dat$date,FUN=sum))
x<-cbind(x,as.data.frame(tapply(dat$steps,dat$date,FUN=mean)))
x<-cbind(x,as.data.frame(tapply(dat$steps,dat$date,FUN=median)))
colnames(x)<-c("Total Steps","Mean","Median")
x
```

## Histogram of total number of steps taken each day
```{r plot}
hist(tapply(dat$steps,dat$date,FUN=sum),main="Total number of steps per day",xaxt='n',xlab="Days")
```

## Mean & Median Number of Steps taken each day
```{r}
mean1<-mean(tapply(dat$steps,dat$date,FUN=sum),na.rm=T)
median1<-median(tapply(dat$steps,dat$date,FUN=sum),na.rm=T)
mean1
median1
```
The mean of the total number of steps taken per day is `r mean1`.  
The median of the total number of steps taken per day is `r median1`.

## What is the average daily activity pattern?
```{r}
time_series<-tapply(dat$steps,dat$interval,mean,na.rm=T)
plot(x=unique(dat$interval),y=time_series,type="l",main="Time Series Plot",xlab="Interval",ylab="Average Number of Steps")
```

### Which 5-minute interval that, on average, contains the maximum number of steps?
```{r}
max_steps<-as.numeric(names(which.max(time_series)))
max_steps
```
The 5-minute interval that contains the maximum number of steps is `r max_steps`.

## Imputing missing values

### Calculate and report the total number of missing values in the dataset
```{r}
missing<-sum(is.na(dat))
missing
```
The number of rows with NA is `r missing`.

### Strategy for filling in the missing values
Fill in missing values with mean for 5-minute interval, the assumption is that the number of steps taken in the 5-minute interval can be well approximated by the mean which is found in the time_series data.  
```{r}
dat1<-dat
for (i in 1:nrow(dat1)){
    if (is.na(dat1[i,1])==TRUE){
        dat1[i,1]<-time_series[paste(dat1$interval[i])]
        }
}
```
dat1 is the new dataset with missing data filled in.  

### Histogram of the total number of steps taken each day with NA values filled in
```{r}
hist(tapply(dat1$step,dat1$date,sum),main="Total number of steps per day",xaxt='n',xlab="Days")
```

### Mean and Median of total number of steps taken per day###
```{r}
mean2<-mean(tapply(dat1$step,dat1$date,sum))
median2<-median(tapply(dat1$step,dat1$date,sum))
mean2
median2
```
The mean `r mean2` and median `r median2` here is about the same as the first part of the assignment. Since we impute data with the mean, the distribution of the histogram stays about the same.

## Are there differences in activity patterns between weekdays and weekends?

### Creating the factor variable
```{r,message=FALSE}
if (!require(timeDate)){
    install.packages("timeDate")
}
library(timeDate)
dat1$day<-ifelse(isWeekend(dat1$date),"Weekend","Weekday")
```

### Differences in activity patterns between weekdays and weekends?

```{r,message=FALSE}
if (!require(lattice)){
    install.packages("lattice")
}
library(lattice)

dat2<-data.frame(interval=unique(dat1$interval),step=tapply(dat1$steps[dat1$day=="Weekday"],dat1$interval[dat1$day=="Weekday"],mean),day="Weekday")
rownames(dat2)<-1:nrow(dat2)

dat3<-data.frame(interval=unique(dat1$interval),step=tapply(dat1$steps[dat1$day=="Weekend"],dat1$interval[dat1$day=="Weekend"],mean),day="Weekend")
rownames(dat3)<-nrow(dat2)+1:nrow(dat2)+nrow(dat3)

dat4<-rbind(dat2,dat3)

xyplot(step~interval|day, data=dat4,layout=c(1,2),type="l",ylab="Number of steps",xlab="Interval")
```

The panel plot that shows the average number of steps for each interval.
